---
title: "01_ManuscriptFigures"
author: "CPJ"
date: "2026-04-01"
output: 
  html_document:
    number_sections: true
---

This script will produce figures for the manuscript "Peatlands Control Aquatic Carbon Pools and Fluxes in an Arctic Subcatchment".

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, 
                      message = FALSE)
```

# Setup

Setting consistent color palettes and installing libraries

```{r Setup, include=FALSE}
####Load in Libraries####
library(dplyr)
library(lubridate)
library(colorblindcheck)
library(ggpubr)
library(rstatix)
library(readr)
library(patchwork)
library(tools)
library(tidyr)
library(randomForest)
library(caret)
library(ComplexUpset)
library(ggrepel)
library(pairwiseAdonis)
library(flextable)
library(officer)

####Import Data Frames####
avg_site_all <- read_csv("avg_site_all.csv")
ft_site_all <- read_csv("ft_site_all.csv")
avg_site_surf <- read_csv("avg_site_surf.csv")

####NMDS Dataframes####
scores_df <- readRDS("nmds_scores_df.rds")
env_vectors_sig_scaled <- readRDS("nmds_env_vectors_scaled.rds")
stress_val <- readRDS("nmds_stress.rds")
r2_pw_df<- read_csv("porewater_perm.csv")
r2_df<- read_csv("all_perm.csv")

###Random Forest Models###
final_rf_pco2 <- readRDS("final_rf_pco2.rds")
final_rf <- readRDS("final_rf.rds")
imputed_data_pco2<-readRDS("imputed_data_pco2.rds")
imputed_data<-readRDS("imputed_data.rds") #co2flux
```

## Plot Styles

Setting up base plotting functions and color palette

```{r Plotting Setup, include=FALSE}
####Set Palettes###
reach_colors<- c (
  "Tundra" = "#9C8C8B", 
  "Forest" = "#8F9C59",
  "Peatland" = "#A25933"
)

reach_shapes<- c (
  "Lake" = 25, 
  "Stream" = 21,
  "Porewater" = 13
)

palette_check(reach_colors, plot=T) #check color blind friendly

####reorder so they plot in the correct order##
avg_site_all$Landcover <- factor(avg_site_all$Landcover,
    levels = c('Tundra','Forest', 'Peatland'),ordered = TRUE)
avg_site_all$Landcover<-as.factor(avg_site_all$Landcover)
ft_site_all$Landcover <- factor(ft_site_all$Landcover,
    levels = c('Tundra','Forest', 'Peatland'),ordered = TRUE)
ft_site_all$Landcover<-as.factor(ft_site_all$Landcover)
avg_site_all$Aquatic_Type <- factor(avg_site_all$Aquatic_Type,
                                levels = c("Stream", "Lake", "Porewater"))
ft_site_all$Aquatic_Type <- factor(ft_site_all$Aquatic_Type,
                                levels = c("Stream", "Lake", "Porewater"))
###make theme####
my_theme<-function() {
  theme_bw() +
  theme(text=element_text(size=16),
        rect=element_rect(fill="transparent"),
        strip.background = element_rect(color="black", fill = "#F2F2F2" , size=1.5 ),
        panel.grid = element_blank(),
        #axis.title.x=element_text(size=30),
        axis.title.x=element_blank(),
        axis.title.y = element_text(size=16), 
        legend.position = "None",
        legend.key = element_rect(colour = NA, fill = NA),
        strip.text=element_text(size=16),
        axis.text.y =element_text(color="black"),
        axis.text.x =element_text(color="black", angle=45, hjust=1, size=16),
        panel.background =element_rect(color="black", fill='transparent', size=1.5),
        plot.background = element_rect(fill = "white", colour = NA),
        plot.margin = ggplot2::margin(10,18,10,18))

}

my_theme2 <- function() {
  theme_bw() +
    theme(
      text = element_text(size = 13),
      plot.background = element_rect(fill = "white", color = NA),
      strip.background = element_rect(color = "black", fill = "#F2F2F2", size = 1.5),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 13),
      legend.position = "right",
      strip.text = element_text(size = 11),
      legend.key = element_rect(colour = NA, fill = NA),
      axis.text.y = element_text(color = "black", size = 13),
      axis.text.x = element_text(color = "black", size = 13),
      panel.background = element_rect(fill = "transparent"),
      panel.grid.major = element_line(color = "grey90"),   # lighter major grids
      panel.grid.minor = element_line(color = "grey95"),   # even lighter minor grids
      panel.border = element_rect(color = "black", fill = NA, size = 1.5), # border on top
      plot.margin = ggplot2::margin(2,3,2,2)
    )
}

#### Base plotting function for boxplots###
plot_flux <- function(df, yvar, ylab_text) {
  # Remove rows where yvar or Aquatic_Type or Landcover are NA
  df_clean <- df %>%
    dplyr::filter(!is.na(.data[[yvar]]), 
                  !is.na(Aquatic_Type), 
                  !is.na(Landcover))
  
  ggplot(df_clean, aes(x = Aquatic_Type, y = .data[[yvar]], fill = Landcover)) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) +
    geom_jitter(
      aes(shape = Aquatic_Type),
      position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
      size = 2.5,
      alpha = 1,
      stroke = 0.7,
      show.legend = FALSE
    ) +
    scale_fill_manual(values = reach_colors, name = "Landcover Type") +
    scale_shape_manual(values = reach_shapes) +
    scale_color_manual(values = reach_colors) +
    ylab(ylab_text) +
    my_theme2() +
    NULL
}

```

# Figure 3- General Chemistry

*Note that stats bar were added after in Powerpoint*

```{r Figure 2- General Chemistry, echo= FALSE}
### DOC (NPOC)####
p1 <- plot_flux(
  avg_site_all,
  "DOC_mgL",
  expression("DOC (mg L"^-1*")")
  ) +
  coord_cartesian(ylim = c(0, 17)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank()
  )

## TDN ####
p2 <- plot_flux(
  avg_site_all,
  "TDN_mgL",
 expression("TDN (mg L"^-1*")")
 ) 


## CN ####
p3 <- plot_flux(
  avg_site_all,
  "CN",
  "DOC:TDN"
) +
  coord_cartesian(ylim = c(0, 40)) 

## CO2 Flux ####
p4 <- plot_flux(
  avg_site_all,
  "CO2Flux_gm2d",
  expression(CO[2]~"Flux"~(g~C~m^-2~d^-1))
) +
  #coord_cartesian(ylim = c(0, 250)) + 
  scale_y_log10(limits = c(0.1, 1500),
                breaks = c(1, 10,20, 50, 100, 200),
  labels = c(1, 10,20, 50, 100, 200)
) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank()
  )

## pCO2 ####
p5 <- plot_flux(
  avg_site_all,
  "pCO2_uatm",
  expression(pCO[2]~"("*mu*"atm)")
) +
  coord_cartesian(ylim = c(0, 15000)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank()
  )

###All####
fig3<-ggarrange (p5, p4, p1, p2, p3, common.legend = TRUE, legend = "top", ncol = 2, nrow = 3, align = "v")
fig3
#ggsave("fig3.png", width=7.5, height=8, units = "in", dpi=600, bg = "white")
```

# Figure 4 - DOM Composition

## Figure 4a- Van krevelan

### Set up

```{r Figure 4a- Plot Set Up, include= FALSE}
#### Van Krevelan Set up ####
##make the rectangles
classification <- tribble(
  ~Class, ~OC_low, ~OC_high, ~HC_low, ~HC_high,
  'Lipid', 0, 0.3, 1.5, 2.5,
  'Unsat. HC', 0, 0.125, 0.8, 1.5,
  'Cond. HC', 0, 0.95, 0.2, 0.8,
  'Protein', 0.3, 0.55, 1.5, 2.3,
  'Amino sugar', 0.55, 0.7, 1.5, 2.2,
  'Carbohydrate', 0.7, 1.5, 1.5, 2.5,
  'Lignin', 0.125, 0.65, 0.8, 1.5,
  'Tannin', 0.65, 1.1, 0.8, 1.5
) %>%
  mutate(label_x = (OC_low + OC_high) / 2,
         label_y = HC_high - 0.1)

## Compound class rectangles ----
class_rect <- geom_rect(
  data = classification,
  aes(
    xmin = OC_low,
    xmax = OC_high,
    ymin = HC_low,
    ymax = HC_high
  ),
  color = 'black',
  fill = NA,
  linewidth = 1,
  inherit.aes = FALSE,
  linetype = 'dashed'
)

rect_label <- geom_label(
  data = classification,
  aes(
    x = label_x,
    y = label_y,
    label = Class
  ),
  inherit.aes = FALSE,
  size = 2.5
)
```

### Data

```{r Figure 4a - Data set up, include= FALSE}
####Now Identify Universal Points####
#Filter out porewater
ft.surf<-ft_site_all %>%
  filter(Sample_Type != "porewater")

# Make sure factors are ordered
ft.surf <- ft.surf %>%
  mutate(
    Aquatic_Type = factor(Aquatic_Type, levels = c("Stream", "Lake")),
    Landcover    = factor(Landcover, levels = c("Tundra", "Forest", "Peatland"))
  )

# ---- Flag universal and partial compounds ----
flags <- ft.surf %>%
  group_by(OC, HC) %>%
  summarize(
    n_group = n_distinct(Landcover),   # now using Landcover 
    total_group = n_distinct(ft.surf$Landcover),
    .groups = "drop"
  ) %>%
  mutate(
    Universal = n_group == total_group,
    Partial   = n_group > 1 & n_group < total_group
  ) %>%
  select(OC, HC, Universal, Partial)

# Join flags back
ft.surf <- ft.surf %>%
  left_join(flags, by = c("OC", "HC"))
```

### Plot

```{r Figure 4- Plot Van Krevelan, echo=FALSE}
van.p <- ggplot(ft.surf, aes(x = OC, y = HC)) +

  # Universal points
  geom_point(
    data = ft.surf %>% filter(Universal == TRUE),
    shape = 21, fill = "#DADEDF", color = "#DADEDF",
    size = 1, alpha = 0.5
  ) +

  # Partial shared points
  geom_point(
    data = ft.surf %>% filter(Partial == TRUE),
    shape = 21, fill = "#ADD8E6", color = "#ADD8E6",
    size = 1, alpha = 0.5
  ) +

  # Unique points (colored by Landcover)
  geom_point(
    data = ft.surf %>% filter(!Universal & !Partial),
    aes(fill = Landcover),
    size = 3, shape = 21, color = "black"
  ) +

  class_rect +
  rect_label +

  scale_fill_manual(values = reach_colors) +

  facet_grid(
    rows = vars(Aquatic_Type),
    cols = vars(Landcover),
    scales = "free"
  ) +

  my_theme() +
  labs(x = "O/C", y = "H/C", fill = "Unique Compounds") +

  theme(
    axis.title.x = element_text(size = 13),
    axis.text.x = element_text(size = 13)
  ) +

  guides(
    fill = guide_legend(
      title.position = "top",
      override.aes = list(size = 10)
    )
  )
van.p

##Now we need to steal the legend from this plot since it's not coming through properly
ggplot(ft.surf, aes(x = OC, y = HC)) +

  # Universal points (grey)
  geom_point(
    data = ft.surf %>% filter(Universal),
    aes(fill = "Universal")   # map to legend
    , color = "#DADEDF",
    size = 3
  ) +

  # Partial points (blue)
  geom_point(
    data = ft.surf %>% filter(Partial),
    aes(fill = "Partial"),     # map to legend
     color = "#ADD8E6",
    size = 3
  ) +

  # Subcategory points (no legend)
  geom_point(
    data = ft.surf %>% filter(!Universal & !Partial),
    aes(x = OC, y = HC), size = 3, color = "black", fill = "white" # adjust color as you like
  ) +

  class_rect +
  rect_label +

  # Only include Universal and Partial in the legend
  scale_fill_manual(
    name = "Common Compounds",
    values = c("Universal" = "#DADEDF", "Partial" = "#ADD8E6")
  ) +

  facet_grid(rows = vars(Aquatic_Type), cols = vars(Landcover), scales = "free") +

  my_theme() +
  labs(x = "O/C", y = "H/C") +

  theme(
    axis.title.x = element_text(size = 13),
    legend.position = "top",
    legend.direction = "horizontal",
    legend.justification = "left",
    legend.box = "horizontal",
    legend.key = element_rect(colour = NA, fill = NA),
    legend.text = element_text(size = 13),
    axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, size = 13)
  ) +
  guides(
    fill = guide_legend(
      title.position = "top",
      override.aes = list(size = 8, color = c("#ADD8E6", "#DADEDF"))
    )
  )
```

##Figure 4b - Upset

###Data

```{r Figure 4b- Data, include=FALSE}
#--------------------------------------------------
# 1. Filter out Porewater and keep unique formulas
#--------------------------------------------------
df.comp_pw <- ft_site_all %>%
  filter(!is.na(Landcover), Aquatic_Type != "Porewater") %>%
  distinct(MolecularFormula, Aquatic_Type, Landcover, Class)

#--------------------------------------------------
# 2. Create presence / absence matrix
#--------------------------------------------------
presence_pw <- df.comp_pw %>%
  mutate(Group = paste(Aquatic_Type, Landcover, sep = "_")) %>%
  select(MolecularFormula, Group) %>%
  distinct() %>%
  mutate(present = 1) %>%
  pivot_wider(
    names_from  = Group,
    values_from = present,
    values_fill = 0
  )

#--------------------------------------------------
# 3. Convert to logical matrix for ComplexUpset
#--------------------------------------------------
upset_data <- presence_pw %>%
  select(-MolecularFormula) %>%
  mutate(across(everything(), ~ . == 1))

#--------------------------------------------------
# 4. Reorder columns: Stream first, Lake second
#--------------------------------------------------
stream_cols <- grep("^Stream", names(upset_data), value = TRUE)
lake_cols   <- grep("^Lake", names(upset_data), value = TRUE)

# Optional: enforce landcover order within each aquatic type
stream_cols <- stream_cols[match(
  c("Stream_Tundra", "Stream_Forest", "Stream_Peatland"),
  stream_cols
)]

lake_cols <- lake_cols[match(
  c("Lake_Tundra", "Lake_Forest", "Lake_Peatland"),
  lake_cols
)]

new_order <- c(stream_cols, lake_cols)
new_order <- new_order[!is.na(new_order)]

upset_data <- upset_data[, new_order]

#--------------------------------------------------
# 5. Clean labels for display
#--------------------------------------------------
colnames(upset_data) <- gsub("_", " ", colnames(upset_data))

```

###Plot

```{r Figure 4b- Upset Plot, echo=FALSE}
up.p <- upset(
  upset_data,
  colnames(upset_data),
  sort_sets = FALSE,

  # control relative panel sizes
  width_ratio = 0.15,   # left bar panel (set sizes)
  height_ratio = 0.6,   # top panel (intersection sizes)

  # base annotations (intersection size bars)
  base_annotations = list(
    "Intersection size" = intersection_size(
  counts = FALSE,
  fill = "steelblue"
) +
  scale_y_continuous(
    breaks = seq(0, 3000, 500)
  ) +
  theme(
    # Pull y-axis label closer to the bars
    axis.title.y = element_text(
      angle = 90,                   # vertical
      vjust = -20,                  # adjust vertical position along panel
      hjust = 0.5,                  # adjust horizontal alignment
      margin = ggplot2::margin(r = 2) # minimal space to bars
    ),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
  ),

  # set sizes (left panel bars)
  set_sizes = upset_set_size(
    geom = geom_bar(fill = "steelblue")
  ) +
    theme(
      # move set size label closer
      axis.title.x = element_text(
        margin = ggplot2::margin(t = 2)
      ),
      # remove y-axis "group" numbers
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
) +
  theme_minimal(base_size = 13) +
  theme(
    # x-axis text for set labels (rotate if long)
    axis.text.x = element_blank(),

    # remove all remaining axis titles
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),

    # tight spacing
    panel.spacing = unit(0.2, "lines"),
    plot.margin = ggplot2::margin(2, 2, 2, 2)
  )

up.p
```

##Figure 4c - Bar Plot

###Data

```{r Figure 4c- Data, include=FALSE}
#--------------------------------------------------
# 1. Remove porewater + define compound groups
#--------------------------------------------------
df_comp <- ft.surf %>%
  mutate(
    Compound_Group = case_when(
      Universal ~ "Universal",
      !Universal & !Partial ~ "Unique",
      TRUE ~ "Shared"
    )
  ) %>%
  filter(Compound_Group %in% c("Universal", "Unique"))

#--------------------------------------------------
# 2. Summarize composition
#--------------------------------------------------
df_plot <- df_comp %>%
  group_by(Aquatic_Type, Landcover, Compound_Group, Class) %>%
  summarise(
    Count = mean(NormIntensity, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  group_by(Aquatic_Type, Landcover, Compound_Group) %>%
  mutate(
    Perc = Count / sum(Count) * 100
  )

#--------------------------------------------------
# 3. Factor ordering
#--------------------------------------------------
df_plot$Landcover <- factor(
  df_plot$Landcover,
  levels = c("Tundra", "Forest", "Peatland")
)

df_plot$Compound_Group <- factor(
  df_plot$Compound_Group,
  levels = c("Universal", "Unique")
)

df_plot$Class <- factor(
  df_plot$Class,
  levels = c(
    "Cond. HC",
    "Tannin",
    "Lignin",
    "Unsat. HC",
    "Protein",
    "Amino sugar",
    "Carbohydrate",
    "Lipid",
    "Other"
  )
)

df_plot$Aquatic_Type <- factor(
  df_plot$Aquatic_Type,
  levels = c("Stream", "Lake")
)

```

###Plot

```{r Figure 4c- Plot, echo=FALSE}
class_colors <- c(
   "Lipid"        = "#E6B89C",  # Soft peach/orange
  "Unsat. HC"    = "#D4C67A",  # Muted yellow-green
  "Cond. HC"     = "#C4886B",  # Warm terracotta
  "Protein"      = "#7BAF91",  # Soft muted teal
  "Amino sugar"  = "#A7C9C2",  # Light teal/blue
  "Carbohydrate" = "#7C9AC8",  # Soft blue
  "Lignin"       = "#B08DA0",  # Dusty pink/lavender
  "Tannin"       = "#9E8F83" ,  # Muted warm gray
  "Other"        = "#B8C0C8"   # Soft gray-blue
)

comp.bar <- ggplot(df_plot, aes(x = Landcover, y = Perc, fill = Class)) +
  geom_col(color = "black") +
  scale_fill_manual(values = class_colors) +
  facet_grid(Compound_Group ~ Aquatic_Type) +
  my_theme2() +
  theme(
    text = element_text(size = 13),
    legend.position = "right",
    legend.key = element_rect(colour = NA, fill = NA),
    axis.title.y = element_blank(),
    axis.title.x = element_text()
  ) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_flip() +
  labs(y = "Percent (%)", x = NULL)

# comp.bar
```

## Figure 4- All

Now put it all together

```{r Figure 4- All, echo=FALSE}
ggarrange(van.p, up.p, comp.bar, nrow=3, heights = c(2, 1, 1), widths = c(1,1,1.3))
ggsave("fig4.png", width=9, height=12, units = "in", dpi=600, bg = "white")
```

# Figure 5- NOSC and Weight

##Data

```{r Data, include=FALSE}
df_mass <- ft.surf %>%
  mutate(
    Compound_Group = case_when(
      Universal ~ "Universal",
      !Universal & !Partial ~ "Unique",
      TRUE ~ "Shared"
    )
  ) %>%
  filter(Compound_Group %in% c("Universal", "Unique"))

# Factor levels
df_mass$Aquatic_Type <- factor(df_mass$Aquatic_Type)
df_mass$Subcategory <- factor(df_mass$Landcover,
                              levels = c("Tundra", "Forest", "Peatland"))
df_mass$Compound_Group <- factor(df_mass$Compound_Group,
                                 levels = c("Universal", "Unique"))
```

##Plots

```{r NOSC Plot, echo=FALSE}
nosc_violin_facet <- ggplot(df_mass, aes(x = Aquatic_Type, y = NOSC, fill = Landcover)) +
  geom_violin(alpha = 0.6, trim = FALSE, position = position_dodge(width = 0.8)) +
  geom_boxplot(width = 0.12, position = position_dodge(width = 0.8),
               outlier.shape = NA, alpha = 0.7) +
  scale_fill_manual(values = reach_colors, name = "Landcover Type") +
  scale_shape_manual(values = reach_shapes) +
  scale_color_manual(values = reach_colors) +
  facet_wrap(~Compound_Group) +
  ylab("NOSC") +
  ylim(-2.5,3) +
  my_theme2() +
  NULL

nosc_violin_facet
```

```{r Mass Plot, echo=FALSE}
mass_violin_facet <- ggplot(df_mass, aes(x = Aquatic_Type, y = Mass, fill = Landcover)) +
  geom_violin(alpha = 0.6, trim = FALSE, position = position_dodge(width = 0.8)) +
  geom_boxplot(width = 0.12, position = position_dodge(width = 0.8),
               outlier.shape = NA, alpha = 0.7) +
  scale_fill_manual(values = reach_colors, name = "Landcover Type") +
  scale_shape_manual(values = reach_shapes) +
  scale_color_manual(values = reach_colors) +
  ylim(0,1350) +
  facet_wrap(~Compound_Group) +  # faceted by Universal / Unique
  ylab("Mass (Da)") +
  my_theme2() +
  NULL

mass_violin_facet
```

```{r Plot save, echo=FALSE}
nm_fac<-ggarrange(nosc_violin_facet, mass_violin_facet, ncol = 1, nrow = 2,common.legend = TRUE,
  legend = "right")
nm_fac

ggsave(
  "fig5.png",
  width = 10,
  height = 10,  
  dpi = 600,
  bg = "white"
)
```

# Figure 6 - DIC

##Stats

```{r Stats for the plot, include=FALSE}
#filter out porewater 
avg_site_surf<-avg_site_all %>%
  filter(Sample_Type != "porewater")

# Fit linear model
model <- lm(del_DIC ~ DistancetoOutlet_m, data = avg_site_surf)
model_summary <- summary(model)

# Extract stats
r2 <- round(model_summary$r.squared, 3)
p_value <- signif(coef(model_summary)[2, 4], 3)

# Optional: format p-value display
p_label <- ifelse(p_value < 0.001, "p < 0.001", paste0("p = ", p_value))
label_text <- paste0("R² = ", r2, "\n", p_label)
```

##Plot

```{r DIC Figure, echo=FALSE}
ggplot(avg_site_surf, aes(x = DistancetoOutlet_m, y = del_DIC)) +
  geom_point(aes(shape = Aquatic_Type, fill = Landcover), 
             size = 3, color = "black", stroke = 0.6) +
  scale_fill_manual(values = reach_colors) +
  scale_shape_manual(values = reach_shapes) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, color = "black") +  # single LM line
  labs(
    x = "Distance from Mouth (m)",
    y = expression(delta^13*C-DIC~"(\u2030)"),
    shape = "Aquatic Environment",
    fill = "Landcover Type"
  ) +
  guides(
  fill = guide_legend(override.aes = list(shape = 21, color = "black", size = 5))
) +
  annotate("text",
           x = 10500,
           y = 0,
           label = label_text,
           hjust = 0, vjust = 1, size = 5) +
  my_theme2() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 13),
    axis.text.y = element_text(size = 13),
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(size = 13),
    axis.title.y  = element_text(size = 13),
    legend.position = "right",
    legend.key = element_rect(colour = NA, fill = NA),
    legend.text = element_text(size = 13)
  ) + 
  scale_x_reverse(limits = c(10500, 0)) +
  scale_y_continuous(limits = c(-20, 0))

ggsave("fig6.png", width=8, height=4, units = "in", dpi=600, bg = "white")
```

# Figure 7 - NMDS

## Plots

###NMDS

```{r NMDS Plot, echo=FALSE}
env_vectors_sig_scaled <- env_vectors_sig_scaled %>%
  mutate(Variable = recode(Variable,
                           "Unsat_HC" = "Unsat. HC",
                           "Cond_HC" = "Cond. HC",
                           "NOSC_mean" = "NOSC",
                           "Shannon_Diversity" = "Shannon Diversity"))

nmdsp<-ggplot(scores_df, aes(x = NMDS1, y = NMDS2)) +
  # Fillable shapes: Stream and Lake
  geom_point(
    data = scores_df %>% filter(Aquatic_Type %in% c("Stream", "Lake")),
    aes(shape = Aquatic_Type, fill = Landcover),
    color = "black",
    size = 3
  ) +
  # Non-fillable shape: Porewater
  geom_point(
    data = scores_df %>% filter(Aquatic_Type == "Porewater"),
    aes(shape = Aquatic_Type, color = Landcover),
    size = 3
  ) +
  # Shape and color scales
  scale_shape_manual(name = "Aquatic Environment", values = reach_shapes) +
  scale_fill_manual(values = reach_colors) +
  scale_color_manual(values = reach_colors) +
  # Ellipses: dotted lines, colored by Landcover
  stat_ellipse(
    aes(group = Landcover, color = Landcover),
    type = "t",
    geom = "path",
    linetype = "dashed",
    fill = NA,
    show.legend = FALSE
  ) +
  # Envfit arrows
  geom_segment(
    data = env_vectors_sig_scaled,
    aes(x = 0, y = 0, xend = MDS1, yend = MDS2),
    arrow = arrow(length = unit(0.3, "cm")),
    color = "black",
    inherit.aes = FALSE
  ) +
  geom_text_repel(
    data = env_vectors_sig_scaled,
    aes(x = MDS1, y = MDS2, label = Variable),
    size = 3.5,
    segment.color = "black",
    inherit.aes = FALSE
  ) +
  annotate(
    "text",
    x = Inf, y = -Inf,
    label = paste("Stress =", stress_val),
    hjust = 1.1, vjust = -0.5,
    size = 5
  ) +
  # Light background grid + panel border
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(),
    axis.title.x = element_text(),
    axis.text.y = element_text(),
    axis.title.y = element_text(),
    legend.position = "right",
    legend.key = element_blank(),
    panel.grid.major = element_line(color = "gray85"),
    panel.grid.minor = element_line(color = "gray90"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.7)
  )
nmdsp
```

###Bar Plots

```{r Plot, echo=FALSE}
r2_pw_plot <- r2_pw_df %>%
  arrange(R2) %>%
  mutate(System = factor(System, levels = System)) %>%
  ggplot(aes(x = System, y = R2)) +
  geom_col(fill = "#7C9AC8") +
  geom_text(aes(label = Signif), hjust = -0.2, size = 6) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Surface:Porewater", y = expression(R^2), x = "") +
  theme_minimal(base_size = 13) +
  theme(
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5),
    panel.border = element_rect(color = "black", fill = NA, size = 0.8)
  )

r2_all_plot <- r2_df %>%
  arrange(R2) %>%
  mutate(Model = factor(Model, levels = Model)) %>%
  ggplot(aes(x = Model, y = R2)) +
  geom_col(fill = "#7C9AC8") +
  geom_text(aes(label = Signif), hjust = -0.2, size = 6) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  coord_flip() +
  labs(title= "Surface", y = expression(R^2), x = "") +
  theme_minimal(base_size = 13) +
  theme(
    axis.title.y = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5),
    panel.border = element_rect(color = "black", fill = NA, size = 0.8)
  )

comb_perm <- ggarrange(
  r2_all_plot, r2_pw_plot,
  ncol = 2
)

comb_perm
```

```{r Save, echo=FALSE}
ggarrange(
  nmdsp, comb_perm,
  ncol = 1,
  nrow = 2,
  #widths = c(1, 2, 1),  # makes the center column wider
  heights = c(2, 1)     # adjust relative heights of top vs bottom
)
ggsave("fig7.png", width=8, height=8, units = "in", dpi=600, bg = "white")
```

# Figure 8 - Random Forest Model

## Figure 8a - pCO2

```{r pCO2, echo=FALSE}
clean_labels <- function(x) {
  x[x == "Raos_Elemental_Comp"] <- "Elemental Diversity"
  x[x == "pCO2_uatm"] <- "pCO2"
  x[x == "DOC_mgL"] <- "DOC"
   x[x == "TDN_mgL"] <- "TDN"
  x <- gsub("_qbased|_mean", "", x)  # remove qbased/mean
   x <- gsub("DistancetoOutlet_m", "Distance to Outlet", x, ignore.case = TRUE)
   x <- gsub("_", " ", x)             # replace underscores with spaces
  x <- gsub("\\.x|\\.y", "", x)      # remove .x/.y
  x <- gsub("CN", "C:N", x, ignore.case = TRUE)
  x[x == "Landcover"] <- "Landcover Type"  # rename Reach_Type
  x <- toTitleCase(x)                 # optional: title case
  x <- gsub("K600", "k600", x, ignore.case = TRUE)
  x
}
# Extract %IncMSE
mse_df_pco2 <- data.frame(
  Variable = rownames(importance(final_rf_pco2)),
  IncMSE = as.numeric(importance(final_rf_pco2)[, "%IncMSE"]),
  stringsAsFactors = FALSE
)

# Sort
mse_df_pco2 <- mse_df_pco2[order(mse_df_pco2$IncMSE, decreasing = TRUE), ]

# Plot
rf_mse_pco2 <- ggplot(mse_df_pco2, aes(x = reorder(Variable, IncMSE), y = IncMSE)) +
  geom_col(fill = "#7C9AC8") +
  coord_flip() +
  scale_x_discrete(labels = clean_labels) +
  labs(
    title = expression("Random Forest on pCO"[2]),
    x = "Variable",
    y = "%IncMSE"
  ) +
  my_theme2() +
  theme(
    axis.title.x = element_text(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
rf_mse_pco2
```

## Figure 8b

### Data

```{r pCO2 Partial Dependence Plots Data, echo=FALSE}
c_pd <- pdp::partial(
  object = final_rf_pco2,
  pred.var = "Raos_Elemental_Comp",
  train = imputed_data_pco2
)



tan_pd <- pdp::partial(
  object = final_rf_pco2,
  pred.var = "Tannin",
  train = imputed_data_pco2
)

n_pd <- pdp::partial(
  object = final_rf_pco2,
  pred.var = "TDN_mgL",
  train = imputed_data_pco2
)
```

### Plots

```{r Partial Dependence Plots, echo=FALSE}
### 1. Get shared y-axis limits


y_min <- min(c(c_pd$yhat,
               tan_pd$yhat,
               n_pd$yhat), na.rm = TRUE)

y_max <- max(c(c_pd$yhat,
               tan_pd$yhat,
               n_pd$yhat), na.rm = TRUE)

### Create individual plots

c_plot <- ggplot(c_pd, aes(x = Raos_Elemental_Comp, y = yhat)) +
  geom_line(size = 1.5, color = "#7C9AC8") +
  labs(x = "Rao's Elemental Diversity", y = NULL) +
  my_theme2() +
  scale_y_continuous(limits = c(y_min, y_max)) +
  theme(axis.title.x = element_text())
  c_plot

tan_plot <- ggplot(tan_pd, aes(x = Tannin, y = yhat)) +
  geom_line(size = 1.5, color = "#7C9AC8") +
  labs(x = "Tannins (Rel Abundance)", y = NULL) +
  my_theme2() +
  scale_y_continuous(limits = c(y_min, y_max)) +
  theme(axis.title.x = element_text())

n_plot <- ggplot(n_pd, aes(x = TDN_mgL, y = yhat)) +
  geom_line(size = 1.5, color = "#7C9AC8") +
  labs(x = expression("TDN (mg L"^-1*")"), y = NULL) +
  my_theme2() +
  scale_y_continuous(limits = c(y_min, y_max)) +
  theme(axis.title.x = element_text())

###Arrange plots

pco2_pp <- ggarrange(
  c_plot, tan_plot, n_plot,
  ncol = 1,
  nrow = 3,
  align = "v"
) 

pco2_pp<-annotate_figure(
  pco2_pp,
  left = text_grob(
    expression(pCO[2]~"("*mu*"atm)"),
    rot = 90,
    size = 13
  )
)
pco2_pp

```

## Figure 8c - CO2 Flux

```{r Random Forest CO2 Flux, echo=FALSE }
# Extract the true %IncMSE column exactly as randomForest uses it
mse_df <- data.frame(
  Variable = rownames(importance(final_rf)),
  IncMSE = as.numeric(importance(final_rf)[, "%IncMSE"]),
  stringsAsFactors = FALSE
)
mse_df
# Sort highest to lowest
mse_df <- mse_df[order(mse_df$IncMSE, decreasing = TRUE), ]

rf_mse_flux <- ggplot(mse_df, aes(x = reorder(Variable, IncMSE), y = IncMSE)) +
  geom_col(fill = "#7C9AC8") +
  coord_flip() +
  scale_x_discrete(labels = clean_labels) +
  labs(
   title = expression("Random Forest on CO"[2]~ "Flux"),
    x = "Variable",
    y = "%IncMSE"
  ) +
  my_theme2() +
  theme(axis.title.x = element_text(),
        panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.title.y = element_blank(),
  plot.title = element_text(hjust = 0.5))

rf_mse_flux
ggsave("CO2fluxrf.png", width=8, height=8, units = "in", dpi=600, bg = "white")
```

## Figure 8d

### Data

```{r CO2 Flux Partial Dependence Plots Data, echo=FALSE}
pd_k <- pdp::partial(
  object = final_rf,
  pred.var = "k600",
  train = imputed_data
)



pd_sl <- pdp::partial(
  object = final_rf,
  pred.var = "Slope",
  train = imputed_data
)

pd_land <- pdp::partial(
  object = final_rf,
  pred.var = "Landcover",
  train = imputed_data
)
```

### Plots

```{r Partial Dependence Plots CO2 Flux, echo=FALSE}
### 1. Get shared y-axis limits


y_min <- min(c(pd_k$yhat,
               pd_sl$yhat,
               pd_land$yhat), na.rm = TRUE)

y_max <- max(c(pd_k$yhat,
               pd_sl$yhat,
               pd_land$yhat), na.rm = TRUE)

### Create individual plots

k_plot <- ggplot(pd_k, aes(x = k600, y = yhat)) +
  geom_line(size = 1.5, color = "#7C9AC8") +
  labs(x = expression(k[600]~"(m d"^{-1}*")"), y = NULL) +
  my_theme2() +
  scale_y_continuous(limits = c(y_min, y_max)) +
  theme(axis.title.x = element_text())
  

s_plot <- ggplot(pd_sl, aes(x = Slope, y = yhat)) +
  geom_line(size = 1.5, color = "#7C9AC8") +
  labs(x = "Slope", y = NULL) +
  my_theme2() +
  scale_y_continuous(limits = c(y_min, y_max)) +
  theme(axis.title.x = element_text())

land_plot <- ggplot(pd_land, aes(x = Landcover, y = yhat)) +
  geom_point(
    size = 6,
    shape = 21,              # allows fill + outline
    fill = "#7C9AC8",        # inside color (your blue)
    color = "black",         # outline color
    stroke = 1             # outline thickness (adjust if needed)
  ) +
  labs(x = "Landcover Type", y = NULL) +
  my_theme2() +
  scale_y_continuous(limits = c(y_min, y_max)) +
  theme(axis.title.x = element_text())

###Arrange plots

co2_pp <- ggarrange(
  k_plot, s_plot, land_plot,
  ncol = 1,
  nrow = 3,
  align = "v"
)

co2_pp<- annotate_figure(
  co2_pp,
  left = text_grob(
    expression(CO[2]~"Flux (g C m"^-2*~"d"^-1*")"),
    rot = 90,
    size = 13
  )
)
co2_pp
```

##Figure 8 all

```{r Figure 8 All, echo=FALSE}
rf_mse_pco2<- rf_mse_pco2 + theme(plot.margin = ggplot2::margin(t = 0, r = 30, b = 0, l = 0))
r<-ggarrange(rf_mse_pco2, pco2_pp, ncol=2)
r<- r +  theme(plot.margin = ggplot2::margin(t = 0, r = 0, b = 15, l = 0))

rf_mse_flux<-rf_mse_flux + theme(plot.margin = ggplot2::margin(t = 0, r = 30, b = 0, l = 0))
r_f<-ggarrange(rf_mse_flux, co2_pp, ncol=2)
r_f<- r_f +  theme(plot.margin = ggplot2::margin(t = 15, r = 0, b = 0, l = 0))

fig8all<-ggarrange(r, r_f, nrow=2)
fig8all

ggsave("Fig8.png", width=10, height=12, units = "in", dpi=600, bg = "white")
```

# Table 1

Table was further tweaked in word for visuals

```{r Table 1 Summary, echo=FALSE}
summary_tbl <- avg_site_surf %>%
  group_by(Aquatic_Type, Landcover) %>%
  summarise(
    n = n(),
    # CO2 flux
    co2flux_mean   = mean(CO2Flux_gm2d, na.rm = TRUE),
    co2flux_median = median(CO2Flux_gm2d, na.rm = TRUE),
    co2flux_sd     = sd(CO2Flux_gm2d, na.rm = TRUE),
    co2flux_min    = min(CO2Flux_gm2d, na.rm = TRUE),
    co2flux_max    = max(CO2Flux_gm2d, na.rm = TRUE),

    # pCO2 (numeric column!)
    pco2_mean   = mean(pCO2_uatm, na.rm = TRUE),
    pco2_median = median(pCO2_uatm, na.rm = TRUE),
    pco2_sd     = sd(pCO2_uatm, na.rm = TRUE),
    pco2_min    = min(pCO2_uatm, na.rm = TRUE),
    pco2_max    = max(pCO2_uatm, na.rm = TRUE),

    .groups = "drop"
  )

summary_stats <- function(x) {
  tibble(
    Mean   = mean(x, na.rm = TRUE),
    Median = median(x, na.rm = TRUE),
    SD     = sd(x, na.rm = TRUE),
    Min    = min(x, na.rm = TRUE),
    Max    = max(x, na.rm = TRUE)
  )
}

# By Reach Type
by_reach <- avg_site_surf %>%
  group_by(Aquatic_Type, Landcover) %>%
  summarise(
    n = n(),
    co2flux = list(summary_stats(CO2Flux_gm2d)),
    pco2    = list(summary_stats(pCO2_uatm)),
    .groups = "drop"
  ) %>%
  unnest(c(co2flux, pco2), names_sep = "_")

# Overall summary
overall <- avg_site_surf %>%
  group_by(Aquatic_Type) %>%
  summarise(
    Landcover = "",
    n = n(),
    co2flux = list(summary_stats(CO2Flux_gm2d)),
    pco2    = list(summary_stats(pCO2_uatm)),
    .groups = "drop"
  ) %>%
  mutate(
    Aquatic_Type = paste("Overall", Aquatic_Type)
  ) %>%
  unnest(c(co2flux, pco2), names_sep = "_")

# Combine
final_table <- bind_rows(by_reach, overall)

final_table <- final_table %>%
  mutate(Blank = " ") %>%   # insert a blank column
  select(Aquatic_Type, Landcover, n,
         co2flux_Mean, co2flux_Median, co2flux_SD, co2flux_Min, co2flux_Max,
         Blank,
         pco2_Mean, pco2_Median, pco2_SD, pco2_Min, pco2_Max)
final_table <- final_table %>%
  mutate(
    across(
      where(is.numeric) & !matches("^n$"),
      ~ round(.x, 2)
    )
  )
last_row <- nrow(final_table)

ft <- flextable(final_table)

# Rename the Aquatic_Type column
ft <- set_header_labels(ft, 
                        Aquatic_Type = "Aquatic Type",
                        Landcover = "Landcover Type",
                        co2flux_Mean   = "Mean",
                        co2flux_Median = "Median",
                        co2flux_SD     = "SD",
                        co2flux_Min    = "Min",
                        co2flux_Max    = "Max",
                        Blank = " ",
                        pco2_Mean      = "Mean",
                        pco2_Median    = "Median",
                        pco2_SD        = "SD",
                        pco2_Min       = "Min",
                        pco2_Max       = "Max"
)

# Add spanner headers
ft <- add_header_row(
  ft,
  values = c("", "", "", "CO2 flux (g m⁻² d⁻¹)", " ", "pCO2 (µatm)"),
  colwidths = c(1, 1, 1, 5, 1, 5)
)

# Center everything
ft <- align(ft, align = "center", part = "all")

# Horizontal line under units (numeric columns only, stop at blank)
ft <- hline(
  ft,
  i = 1,                    # spanner row
  j = c(3:7, 9:13),         # numeric columns only
  border = fp_border(width = 1),
  part = "header"
)

# Bold Overall row
ft <- bold(ft, i = last_row, bold = TRUE)

# Thick line above Overall row
ft <- hline(
  ft,
  i = last_row,
  border = fp_border(width = 2),
  part = "body"
)

# Auto-adjust widths
ft <- autofit(ft)

ft
doc <- read_docx() %>%
  body_add_flextable(ft)  # 'ft' is your formatted flextable

# Save it
print(doc, target = "CO2_pCO2_summary.docx")

#write.csv(final_table, "CO2_pCO2_summary.csv", row.names = FALSE)
```
