---
title: "Supplementary"
author: "CPJ"
date: "2026-04-10"
output: html_document
---
This script will produce supplementary material for the manuscript "Peatlands Control Aquatic Carbon Pools and Fluxes in an Arctic Subcatchment".
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r Setup, include=FALSE}
####Load in Libraries####
library(dplyr)
library(lubridate)
library(colorblindcheck)
library(ggpubr)
library(rstatix)
library(readr)
library(patchwork)
library(tools)
library(tidyr)
library(randomForest)
library(caret)
library(ComplexUpset)
library(ggrepel)
library(pairwiseAdonis)
library(flextable)
library(officer)

####Import Data Frames####
avg_site_all <- read_csv("avg_site_all.csv")
ft_site_all <- read_csv("ft_site_all.csv")

####NMDS Dataframes####
scores_df <- readRDS("nmds_scores_df.rds")
env_vectors_sig_scaled <- readRDS("nmds_env_vectors_scaled.rds")
stress_val <- readRDS("nmds_stress.rds")
r2_pw_df<- read_csv("porewater_perm.csv")
r2_df<- read_csv("all_perm.csv")

###Random Forest Models###
final_rf_pco2 <- readRDS("final_rf_pco2.rds")
final_rf <- readRDS("final_rf.rds")
imputed_data_pco2<-readRDS("imputed_data_pco2.rds")
imputed_data<-readRDS("imputed_data.rds") #co2flux

###Porewater data###
avg_site_all <- avg_site_all %>%
  group_by(Plot) %>%
  mutate(
    Aquatic_Type_pair = first(Aquatic_Type[Sample_Type != "porewater"])
  ) %>%
  ungroup()

avg_site_pw <- avg_site_all %>%
  filter(Sample_Type == "porewater")
avg_site_pw$Aquatic_Type_pair <- factor(avg_site_pw$Aquatic_Type_pair,
                                levels = c("Stream", "Lake"))
avg_site_pw$Landcover <- factor(avg_site_pw$Landcover,
    levels = c('Tundra','Forest', 'Peatland'),ordered = TRUE)
```

## Plot Styles

Setting up base plotting functions and color palette

```{r Plotting Setup, include=FALSE}
####Set Palettes###
reach_colors<- c (
  "Tundra" = "#9C8C8B", 
  "Forest" = "#8F9C59",
  "Peatland" = "#A25933"
)

reach_shapes<- c (
  "Lake" = 25, 
  "Stream" = 21,
  "Porewater" = 13
)

palette_check(reach_colors, plot=T) #check color blind friendly

####reorder so they plot in the correct order##
avg_site_all$Landcover <- factor(avg_site_all$Landcover,
    levels = c('Tundra','Forest', 'Peatland'),ordered = TRUE)
avg_site_all$Landcover<-as.factor(avg_site_all$Landcover)
ft_site_all$Landcover <- factor(ft_site_all$Landcover,
    levels = c('Tundra','Forest', 'Peatland'),ordered = TRUE)
ft_site_all$Landcover<-as.factor(ft_site_all$Landcover)
avg_site_all$Aquatic_Type <- factor(avg_site_all$Aquatic_Type,
                                levels = c("Stream", "Lake", "Porewater"))
ft_site_all$Aquatic_Type <- factor(ft_site_all$Aquatic_Type,
                                levels = c("Stream", "Lake", "Porewater"))
###make theme####
my_theme<-function() {
  theme_bw() +
  theme(text=element_text(size=16),
        rect=element_rect(fill="transparent"),
        strip.background = element_rect(color="black", fill = "#F2F2F2" , size=1.5 ),
        panel.grid = element_blank(),
        #axis.title.x=element_text(size=30),
        axis.title.x=element_blank(),
        axis.title.y = element_text(size=16), 
        legend.position = "None",
        legend.key = element_rect(colour = NA, fill = NA),
        strip.text=element_text(size=16),
        axis.text.y =element_text(color="black"),
        axis.text.x =element_text(color="black", angle=45, hjust=1, size=16),
        panel.background =element_rect(color="black", fill='transparent', size=1.5),
        plot.background = element_rect(fill = "white", colour = NA),
        plot.margin = ggplot2::margin(10,18,10,18))

}

my_theme2 <- function() {
  theme_bw() +
    theme(
      text = element_text(size = 13),
      plot.background = element_rect(fill = "white", color = NA),
      strip.background = element_rect(color = "black", fill = "#F2F2F2", size = 1.5),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 13),
      legend.position = "right",
      strip.text = element_text(size = 11),
      legend.key = element_rect(colour = NA, fill = NA),
      axis.text.y = element_text(color = "black", size = 13),
      axis.text.x = element_text(color = "black", size = 13),
      panel.background = element_rect(fill = "transparent"),
      panel.grid.major = element_line(color = "grey90"),   # lighter major grids
      panel.grid.minor = element_line(color = "grey95"),   # even lighter minor grids
      panel.border = element_rect(color = "black", fill = NA, size = 1.5), # border on top
      plot.margin = ggplot2::margin(2,3,2,2)
    )
}

#### Base plotting function for boxplots###
plot_flux <- function(df, yvar, ylab_text) {
  # Remove rows where yvar or Aquatic_Type or Landcover are NA
  df_clean <- df %>%
    dplyr::filter(!is.na(.data[[yvar]]), 
                  !is.na(Aquatic_Type), 
                  !is.na(Landcover))
  
  ggplot(df_clean, aes(x = Aquatic_Type, y = .data[[yvar]], fill = Landcover)) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) +
    geom_jitter(
      aes(shape = Aquatic_Type),
      position = position_jitterdodge(jitter.width = 0.35, dodge.width = 0.8),
      size = 2.5,
      alpha = 1,
      stroke = 0.7,
      show.legend = FALSE
    ) +
    scale_fill_manual(values = reach_colors, name = "Landcover Type") +
    scale_shape_manual(values = reach_shapes) +
    scale_color_manual(values = reach_colors) +
    ylab(ylab_text) +
    my_theme2() +
    NULL
}

plot_flux_pw <- function(df, yvar, ylab_text) {

  ggplot(df, aes(x = Aquatic_Type, y = .data[[yvar]], fill = Landcover)) +
    
    geom_boxplot(
      aes(group = Landcover),
      alpha = 0.5,
      outlier.shape = NA,
      na.rm = TRUE
    ) +
    
    geom_jitter(
      aes(
        shape = Aquatic_Type_pair,
        group = Landcover
      ),
      position = position_jitterdodge(
        jitter.width = 0.35,
        dodge.width = 0.8
      ),
      size = 2.5,
      alpha = 1,
      stroke = 0.7,
      na.rm = TRUE
    ) +
    
    scale_fill_manual(
      values = reach_colors,
      name = "Landcover Type",
      guide = guide_legend(
        override.aes = list(shape = NA)  # <- removes shapes from fill legend
      )
    ) +
    
    scale_shape_manual(
      values = reach_shapes,
      name = "Aquatic Type"
    ) +
    
    labs(shape = "Aquatic Type") +
    
    ylab(ylab_text) +
    
    my_theme2()
}
```

# Figure S1
## Plot
```{r Porewater}
s1 <- plot_flux_pw(
  avg_site_pw,
  "conc_CO2_mM",
  expression(CO[2]~"Concentration (mM)")
) 
s1
ggsave("figs1.png", width=5, height=3.5, units = "in", dpi=600, bg = "white")
```

## Stats
```{r}
shapiro.test(avg_site_pw$conc_CO2_mM)
kruskal.test(conc_CO2_mM ~ Landcover, data = avg_site_pw)
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
pch4f <- plot_flux(
  avg_site_surf,
  "conc_CH4_mM",
  expression(CH[4]~ "Concentration (mM)")
) 
 
pch4f
ggsave("figs2.png", width=7, height=4, units = "in", dpi=600, bg = "white")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
shandiv <- plot_flux(
  avg_site_surf,
  "Shannon_Diversity",
  expression("Shannon's Diversity (H')")
) 
shandiv

rdiv<-plot_flux(avg_site_surf,
                 "Raos_Elemental_Comp",
                 expression("Rao's Elemental Diversity"))
rdiv

div<-ggarrange(shandiv, rdiv, nrow=2, common.legend = T, legend = "right")
div
ggsave(
  "Suppdiversity.png",
  div,
  width = 7,
  height = 8,   # try 9–10
  dpi = 600,
  bg = "white"
)
```

